---
title: "Day 9 Computational Session"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(reprex)
```

### Warm up 

Starting from penguins, make a new subset in which you:

1.  only keep columns species, island, sex, body_mass_g and flipper_length_mm
2.  filter to only include chinstrap and gentoos
3.  rename the species column to penguin_species
4.  make a summary table that contains the mean of body mass, mean of flipper length and the total count of observations by species and sex

```{r}
penguins_subset <- penguins %>% 
  select(species,island, sex, body_mass_g, flipper_length_mm) %>% 
  filter(species %in% c("Chinstrap","Gentoo")) %>% 
  rename(penguin_species = species) %>% 
  group_by(penguin_species, sex) %>% 
  summarize(mean_body_mass = mean(body_mass_g, na.rm = TRUE),
            mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            total_count = n())
```

A view variations on dplyr::across()

Make a summary table of mean values by species for all numeric variables in the penguins dataset.

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), max , na.rm = TRUE))
```

Calculate the mean value of all columns that start with 'bill'

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(across(starts_with("bill"), min, na.rm= TRUE))
```

Write code that will return the minimum value of any column that end with the character mm, grouped by island and year

```{r}
penguins %>% 
  group_by(island, year) %>% 
  summarize(across(ends_with("mm"),min, na.rm = TRUE))
```

Find and return both the mean and sd of all columns starting with "bill", grouped by penguin species, and returning a new name that combines the original column with the function

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"), 
                   list(mean= mean, sd= sd), 
                   .names = "{.col}.{.fn}", na.rm = TRUE))

```

```{r}
penguins_table <-
  penguins %>% 
  group_by(species, sex) %>% 
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE),
            sd_mass = sd(body_mass_g, na.rm =TRUE)) %>% 
  kable(col.names = c("Species","Sex","Mean body mass (g)","SD body mass(g)")) %>% 
  kable_styling(full_width = FALSE)

penguins_table
```

```{r}
chart1 <- penguins %>% 
  ggplot(aes(x= body_mass_g , y= flipper_length_mm))+
  geom_point()
```

```{r}
chart2 <- penguins %>% 
  ggplot(aes(x = bill_length_mm))+
  geom_histogram()
```

```{r}
(chart1 + chart2)/chart2
```

```{r}
#how to add alt text in a figure caption
#%>%  fig-cap : Penguin body mass observations by species.
#%>%  fig-alt: Jitterplot showing body mass measurements (y-axis, in grams) for three penguin species (Adelie, Chinstrap, and Gentoo) observed near Palmer Archipelago, Antarctica. Adelie and Chinstraps have similar body mass distributions, with Gentoos noticeably larder than both.
ggplot(data= penguins, aes(x= species , y = body_mass_g)) +
  geom_jitter(width = 0.2, alpha = 0.5, color = "navy")
```

Reprex- reproducible examples

```{r}
library(tidyverse)

ggplot(data = mtcars, aes(x= mpg , y = hp))+
  geom_point()
```

```{r}
x = 5 + NA
```
